Can we explain the popularity of Netflix Shows?

### 2023-01-23 submission.Rmd
save work

```{r echo=FALSE, out.width = '100%'}
knitr::include_graphics("netflixHomepage.JPG")
```

@@

# Packages
Before we get started. We will load all the relevant Pacakages
```{r load packages, message=FALSE}
library(reticulate)
library(DBI)
library(dplyr)
library(lubridate)
library(tidyr)
library(data.table)
library(quanteda)
library(tm)
library(lsa)
library("quanteda.textmodels")
library(ggplot2)
library(kableExtra)
library(Hmisc)
library(ggcorrplot)
```

@@

### Source 1: Kaggle 
This dataset provides us with a list of Netflix Shows along with information such as the year it was added, duration, cast, description etc..
There are altogether 8807 Netflix Shows in this dataset.
```{r}
netflixData <- read.csv(file = 'datasets/dataset_main.csv')
```
```{r echo=FALSE}
# FOR Displaying results
# temp = head(netflixData, 10)
# temp %>%
#   kbl() %>%
#   kable_styling() %>%
#   column_spec(1, width = "10cm") %>%
#   scroll_box(width = "1000px", height = "500px")
```

@@

# Cleaning the data

## Removing duplicates or irrelevant observations
## Dealing with duplicates 
Checking if there are any duplicated Netflix Shows
``` {r}
print(length(unique(df$title)))
```

@@

## Handle Missing data
``` {r}
#desc <- describe(df)
#print(desc)
```

```{r}
feature <- new_names # retrieving all the column names

@@

print(NA_Count)

@@

### 2. director
``` {r}
view <- df %>% group_by(director) %>% summarise(count = n())
print(view)
```
Some of the rows have more than one director. 
Action: In the next Feature Engineering Section, I will make mapping tables and use feature engineering to capture this variable

@@

``` {r}
df$country <- sub("\\,.*", "", df$country)
```

### 5. Date Added
``` {r}
view <- df %>% group_by(date_added) %>% summarise(count = n())
print(view)
print(typeof(df$date_added))
print(df$date_added)

```
Date Added is in the character format.
Action: Change this to date format
``` {r}
#df$date_added <- mdy(df$date_added)
#print(df$date_added )
```

@@

Remove string from duration and convert character to integer
``` {r}
df$duration_min <-  gsub('min','',df$duration_min)
df$duration_min <- as.numeric(df$duration_min)
df$duration_season <-  gsub('Seasons','',df$duration_season)
df$duration_season <-  gsub('Season','',df$duration_season)
df$duration_season <- as.numeric(df$duration_season)
print(head(df[,c("title","duration_min","duration_season")]))
```

Drop duration column
```{r}
df = subset(df, select = -c(duration))
```

@@

## Mapping Tables 
As talked about in the Data Cleaning Section above, I will build a mapping table for Cast, Director, Country and Genre. 
This would help reflect their one to many relationship.

@@

For each Show/Movie, take the points of the most famous Actor/Actress and attach it to the main table
```{r}
dbWriteTable(con, "main", df)

query <- "SELECT a.*
        , CASE WHEN c.Points IS NOT NULL THEN max(c.Points)
        ELSE 0 END AS cast_points
        FROM main a 
        LEFT JOIN castPointsMapping c ON a.id = c.id 
        GROUP BY a.id"

@@

query <- "SELECT a.*
        , CASE WHEN d.Points IS NOT NULL THEN max(d.Points)
        ELSE 0 END AS director_points
        FROM main_2 a 
        LEFT JOIN directorPointsMapping d ON a.id = d.id 
        GROUP BY a.id"

@@

Before carrying out further analysis, I disconnect from SQLite I write the main dataframe into a csv 
``` {r}
dbDisconnect(con)
print(head(df,10))

df = subset(df, select = -c(cast,director) )

write.csv(df, "datasets/view_main_cleaned.csv", row.names=FALSE)
```

# Feature Engineering -  Latent Semantic Analysis of Plot and google plot data
# load data

#``` {r}
main <- read.csv(file = 'datasets/view_main_features.csv')
main$plot_combined <- paste(main$plot,main$google_plot)
#```
# Convert data to corpus
#``` {r}
corpus <- Corpus(VectorSource(main$plot_combined))
#```

# Remove special characters, case, punctuation and numbers
#``` {r}

@@

# Remove stop words
#``` {r}
corpus <- tm_map(corpus, removeWords, stopwords("english"))
#```

# Remove white space
#``` {r}
corpus <- tm_map(corpus, stripWhitespace)
#```

# Stemming
#``` {r}
corpus <- tm_map(corpus, stemDocument)
#```

# Term document matrix
#``` {r}
tdm <- TermDocumentMatrix(corpus)
#```

# View top 5 most frequently occuring term
#``` {r}
tdmMatrix <- as.matrix(tdm)
v <- sort(rowSums(tdmMatrix), decreasing=TRUE)
d <- data.frame(word=names(v),freq=v)
print(head(d,5))
#```

# Weighting 
#``` {r}
tdm2 <- as.textmatrix(as.matrix(tdm))
tdm.weighted <- lw_logtf(tdm2)
print(tdm.weighted)
#```

# Latent Semantic Analysis
#``` {r}
lsa <- lsa(tdm.weighted, dims=10)

@@

# Varimax Rotation
#``` {r}
Tk.varimax <- varimax(Tk)

@@

# Interreting Term and Document Matrices
#``` {r}

@@

Dk.thresholded <- as.data.table(threshold.loadings.table(sort.loadings.table(Dk.rotated),q=1))
# print(Dk.thresholded[,.(item,max.factor,max.value)])
write.csv(Dk.thresholded[,.(item,max.factor,max.value)],"datasets/dk.csv", row.names=FALSE)
#```
#``` {r}
biplot <- function(Tk,Dk,xdim,ydim,Tk.scale=1,...){
    print("made it")
    Tk[,type:="Terms"]
    Dk[,type:="Title"]
    biplot.data <- rbindlist(list(Tk,Dk))
    setnames(biplot.data,xdim,"F1")
    setnames(biplot.data,ydim,"F2")
    biplot.data[type=="Terms",F1:=F1*Tk.scale]
    biplot.data[type=="Terms",F2:=F2*Tk.scale]
    ggplot(biplot.data, aes(x=F1, y=F2, label=item, col=type)) + geom_text()
}

@@

#``` {r}

con <- dbConnect(RSQLite::SQLite(), ":memory:")

main <- read.csv(file = 'datasets/view_main_features.csv')
lsa <- read.csv(file = 'datasets/lsa_2.csv')

lsa$item <- sub("^", "s", lsa$item )

@@

query <- "SELECT m.*, l.*
                FROM main m
                LEFT JOIN lsa l ON m.id = l.item"
res <- dbSendQuery(con, query)
df <- dbFetch(res)
dbClearResult(res)
setDT(df)
View(df)

@@

dbDisconnect(con)
#```

@@

1. Type
```{r}
ggplot(df, aes(x = type)) +  # Plot with values on top
  geom_bar() 
plot()
```
2. LSA

3. Date Added
```{r}
ggplot(df, aes(x = date_added)) +
    geom_density()
plot()
```

>>>>

@@

# Packages
Before we get started. We will load all the relevant Pacakages
```{r load packages, message=FALSE}
library(reticulate)
library(DBI)
library(dplyr)
library(lubridate)
library(tidyr)
library(data.table)
library(quanteda)
library(tm)
library(lsa)
library("quanteda.textmodels")
library(ggplot2)
library(kableExtra)
library(Hmisc)
library(ggcorrplot)
library(forcats)
```

@@

### Source 1: Kaggle 
This dataset provides us with a list of Netflix Shows along with information such as the year it was added, duration, cast, description etc..
There are altogether 8807 Netflix Shows in this dataset.
```{r}
netflixData <- read.csv(file = 'datasets/dataset_main.csv')
```

@@

# Cleaning the data
## Removing duplicates or irrelevant observations
### Dealing with duplicates 
Checking if there are any duplicated Netflix Shows
``` {r}
print(length(unique(df$title)))
```

@@

### Handle Missing data
Lets see what data is missing:
```{r}
feature <- new_names # retrieving all the column names

@@

print(NA_Count)
```

@@

### 2. Director
``` {r}
view <- df %>% group_by(director) %>% summarise(count = n())
print(view)
```
Some of the rows have more than one director. 
Action: In the next Feature Engineering Section, I will make mapping tables and use feature engineering to capture this variable

@@

``` {r}
df$country <- sub("\\,.*", "", df$country)

view <- df %>% group_by(country) %>% summarise(count = n())
print(view)
```
Some Countries appear very few times 
Action: I will group them into an "Others" category if count is less than 50
``` {r}
df <- df %>% 
    group_by(country) %>% 
    mutate(country = ifelse(n()>50, country, 'Others'))
```

### 5. Date Added
``` {r}
view <- df %>% group_by(date_added) %>% summarise(count = n())
print(view)
```
Date Added looks good
Action: None

@@

Remove string from duration and convert character to integer
``` {r}
df$duration_min <-  gsub('min','',df$duration_min)
df$duration_min <- as.numeric(df$duration_min)
df$duration_season <-  gsub('Seasons','',df$duration_season)
df$duration_season <-  gsub('Season','',df$duration_season)
df$duration_season <- as.numeric(df$duration_season)
print(head(df[,c("title","duration_min","duration_season")]))
```
Drop duration column
```{r}
df = subset(df, select = -c(duration))
```

@@

## Mapping Tables 
As talked about in the Data Cleaning Section above, I will build a mapping table for Cast and Director

This would help reflect their one to many relationship.

@@

For each Show/Movie, if there are points attached to at least one of the actors/actresses, I give the new variable cast_points the value "known", otherwise, it is labelled as "unknown".
```{r}
dbWriteTable(con, "main", df)

query <- "SELECT a.*
        , CASE WHEN c.Points IS NOT NULL THEN 'known'
        ELSE 'unknown' END AS cast_points
        FROM main a 
        LEFT JOIN castPointsMapping c ON a.id = c.id 
        GROUP BY a.id"

@@

query <- "SELECT a.*
        , CASE WHEN d.Points IS NOT NULL THEN 'known'
        ELSE 'unknown' END AS director_points
        FROM main_2 a 
        LEFT JOIN directorPointsMapping d ON a.id = d.id 
        GROUP BY a.id"

@@

Drop Director and Cast Column
```{r}
df = subset(df, select = -c(cast,director) )
```

Before carrying out further analysis, I disconnect from SQLite I write the main dataframe into a csv 
``` {r}
dbDisconnect(con)
write.csv(df, "datasets/view_main_cleaned.csv", row.names=FALSE)
```

## Data Preview
Here is what the dataset looks like now after cleaning
``` {r echo=FALSE}
temp = head(df, 10)
temp %>%
  kbl() %>%
  kable_styling() %>%
  column_spec(1, width = "10cm") %>%
  scroll_box(width = "1000px", height = "500px")
```

# Feature Engineering -  Latent Semantic Analysis of Plot and google plot data
#### load data

#``` {r}
main <- read.csv(file = 'datasets/view_main_features.csv')
main$plot_combined <- paste(main$plot,main$google_plot)
#```
#### Convert data to corpus
#``` {r}
corpus <- Corpus(VectorSource(main$plot_combined))
#```

#### Remove special characters, case, punctuation and numbers
#``` {r}

@@

#### Remove stop words
#``` {r}
corpus <- tm_map(corpus, removeWords, stopwords("english"))
#```

#### Remove white space
#``` {r}
corpus <- tm_map(corpus, stripWhitespace)
#```

#### Stemming
#``` {r}
corpus <- tm_map(corpus, stemDocument)
#```

#### Term document matrix
#``` {r}
tdm <- TermDocumentMatrix(corpus)
#```

#### View top 5 most frequently occuring term
#``` {r}
tdmMatrix <- as.matrix(tdm)
v <- sort(rowSums(tdmMatrix), decreasing=TRUE)
d <- data.frame(word=names(v),freq=v)
print(head(d,5))
#```

#### Weighting 
#``` {r}
tdm2 <- as.textmatrix(as.matrix(tdm))
tdm.weighted <- lw_logtf(tdm2)
print(tdm.weighted)
#```

#### Latent Semantic Analysis
#``` {r}
lsa <- lsa(tdm.weighted, dims=10)

@@

#### Varimax Rotation
#``` {r}
Tk.varimax <- varimax(Tk)

@@

#### Interreting Term and Document Matrices
#``` {r}

@@

Dk.thresholded <- as.data.table(threshold.loadings.table(sort.loadings.table(Dk.rotated),q=1))
print(Dk.thresholded[,.(item,max.factor,max.value)])
write.csv(Dk.thresholded[,.(item,max.factor,max.value)],"datasets/dk.csv", row.names=FALSE)
#```
###biplot
#``` {r}
biplot <- function(Tk,Dk,xdim,ydim,Tk.scale=1,...){
    print("made it")
    Tk[,type:="Terms"]
    Dk[,type:="Title"]
    biplot.data <- rbindlist(list(Tk,Dk))
    setnames(biplot.data,xdim,"F1")
    setnames(biplot.data,ydim,"F2")
    biplot.data[type=="Terms",F1:=F1*Tk.scale]
    biplot.data[type=="Terms",F2:=F2*Tk.scale]
    ggplot(biplot.data, aes(x=F1, y=F2, label=item, col=type)) + geom_text()
}

@@

``` {r}
con <- dbConnect(RSQLite::SQLite(), ":memory:")

main <- read.csv(file = 'datasets/view_main_cleaned.csv')
lsa <- read.csv(file = 'datasets/lsa_2.csv')

lsa$item <- sub("^", "s", lsa$item)

@@

query <- "SELECT m.*, l.[max.factor] as lsa_value
                FROM main m
                LEFT JOIN lsa l ON m.id = l.item"
res <- dbSendQuery(con, query)
df <- dbFetch(res)
dbClearResult(res)

df = subset(df, select = -c(plot,google_plot))

@@

dbDisconnect(con)
```
# Data Preview 
``` {r echo=FALSE}
temp = head(df, 10)
temp %>%
  kbl() %>%
  kable_styling() %>%
  column_spec(1, width = "10cm") %>%
  scroll_box(width = "1000px", height = "500px")
```

@@

```{r echo=FALSE, out.width = '50%'}
knitr::include_graphics("Type Distribution.JPG")
```
```{r echo=FALSE, out.width = '50%'}
knitr::include_graphics("DirectorDist.JPG")
```
```{r echo=FALSE, out.width = '50%'}
knitr::include_graphics("CastDist.JPG")
```
```{r echo=FALSE, out.width = '50%'}
knitr::include_graphics("DateAddedDist.JPG")
```

<iframe src=" https://public.tableau.com/views/DistributionNetflix/TypeDist?:language=en-US&publish=yes&:display_count=n&:origin=viz_share_link:showVizHome=no&:embed=true"
 width="645" height="955"></iframe>
